---
title: 'List 04: Prediction'
author: "Nikita V. Artamonov"
output:
  pdf_document:
    toc: yes
    df_print: default
    latex_engine: xelatex
  html_document:
    toc: yes
    toc_float:
      collapsed: no
    number_sections: yes
    df_print: paged
  word_document:
    toc: yes
---

```{r, message=FALSE, echo=FALSE}
library(stargazer)
library(lmtest)
library(car)
library(ggplot2)
data(sleep75, wage1, wage2, Labour, Electricity, package = c('wooldridge', 'Ecdat') )
sleep75$gender <- factor(sleep75$male, labels = c('female', 'male') )
wage1$gender <- factor(wage1$female, labels = c('male', 'female') )
```

# sleep equation #1

For the dataset `sleep75` consider a regression
**sleep на totwrk, age, south, male, smsa, yngkid, marr**.

Fitting results:

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.digits.output <- 3
my.digits <- 2
regr <- lm(formula=sleep~totwrk+age+south+male+smsa+yngkid+marr, data=sleep75)
stargazer(regr, type='text', digits=my.digits.output, digit.separator = '', df=FALSE)
```

Consider individuals with the following characteristics:

```{r, echo=FALSE, comment=''}
new.df <- data.frame(totwrk=c(2150, 1950, 2240), age=c(37, 28, 26), 
                     south=c(0, 1, 0), male=c(1,1, 0), smsa=c(1,0,1),
                     yngkid=c(0,1,0), marr=c(1,0,0) )
stargazer(new.df, summary = FALSE, type = 'text', digit.separator = '')
```

Evaluate the prediction of the dependent variable for each individual.  
**Round the answer to `r my.digits` decimal places.**

The answer

```{r, echo=FALSE, comment=''}
my.predict <- predict(regr, new.df)
# names(my.predict) <- 1:nrow(new.df)
# round(my.predict, digits = my.digits )
stargazer(round(data.frame(Prediction=my.predict), digits = my.digits ), 
          type='text', summary=FALSE, digit.separator = '', digits=my.digits)
```

# sleep equation #2

For the dataset `sleep75` consider a regression
**sleep на totwrk, age, age^2^, south, male**.

Fitting results:

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.digits.output <- 3
my.digits <- 2
regr <- lm(formula=sleep~totwrk+age+I(age^2)+south+male, data=sleep75)
stargazer(regr, type='text', digits=my.digits.output, digit.separator = '', df=FALSE)
```

Consider individuals with the following characteristics:

```{r, echo=FALSE, comment=''}
new.df <- data.frame(totwrk=c(2160, 1720, 2390), age=c(32, 24, 44), 
                     south=c(1, 0, 0), male=c(0,1, 1) )
stargazer(new.df, summary = FALSE, type = 'text', digit.separator = '')
```

Evaluate the prediction of the dependent variable for each individual.  
**Round the answer to `r my.digits` decimal places.**

The answer

```{r, echo=FALSE, comment=''}
my.predict <- predict(regr, new.df)
# names(my.predict) <- 1:nrow(new.df)
# round(my.predict, digits = my.digits )
stargazer(round(data.frame(Prediction=my.predict), digits = my.digits ), 
          type='text', summary=FALSE, digit.separator = '', digits=my.digits)
```

# wage equation #1

For the dataset `wage2` consider a regression
**log(wage) на age, IQ, south, married, urban**.

Fitting results:

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.digits.output <- 3
my.digits <- 2
regr <- lm(formula=log(wage)~age+IQ+south+married+urban, data=wage2)
stargazer(regr, type='text', digits=my.digits.output, digit.separator = '', df=FALSE)
```

Consider individuals with the following characteristics:

```{r, echo=FALSE, comment=''}
new.df <- data.frame(age=c(36, 29, 25), IQ=c(105, 123, 112),
                     south=c(1, 0, 1), married=c(1,1,0), urban=c(1,0,1) )
stargazer(new.df, summary = FALSE, type = 'text', digit.separator = '')
```

Evaluate the prediction of the dependent variable for each individual.  
**Round the answer to `r my.digits` decimal places.**

The answer

```{r, echo=FALSE, comment=''}
my.predict <- exp( predict(regr, new.df) )
# names(my.predict) <- 1:nrow(new.df)
# round(my.predict, digits = my.digits )
stargazer(round(data.frame(Prediction=my.predict), digits = my.digits ), 
          type='text', summary=FALSE, digit.separator = '', digits=my.digits)
```

# wage equation #2

For the dataset `wage1` consider a regression
**log(wage) на exper, exper^2^, female, married, smsa**.

Fitting results:

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.digits.output <- 3
my.digits <- 2
regr <- lm(formula=log(wage)~exper+I(exper^2)+female+married+smsa, data=wage1)
stargazer(regr, type='text', digits=my.digits.output, digit.separator = '', df=FALSE)
```

Consider individuals with the following characteristics:

```{r, echo=FALSE, comment=''}
new.df <- data.frame(exper=c(5, 26, 38), female=c(1, 0, 1),
                     married=c(1, 0, 1), smsa=c(1,1,0) )
stargazer(new.df, summary = FALSE, type = 'text', digit.separator = '')
```

Evaluate the prediction of the dependent variable for each individual.  
**Round the answer to `r my.digits` decimal places.**

The answer

```{r, echo=FALSE, comment=''}
my.predict <- exp( predict(regr, new.df) )
# names(my.predict) <- 1:nrow(new.df)
# round(my.predict, digits = my.digits )
stargazer(round(data.frame(Prediction=my.predict), digits = my.digits ), 
          type='text', summary=FALSE, digit.separator = '', digits=my.digits)
```

# output equation #1

For the dataset `Labour` consider a regression
**log(output) на log(capital), log(labour)**.

Fitting results:

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.digits.output <- 3
my.digits <- 2
regr <- lm(formula=log(output)~log(capital)+log(labour), data=Labour)
stargazer(regr, type='text', digits=my.digits.output, digit.separator = '', df=FALSE)
```

Consider firms with the following characteristics:

```{r, echo=FALSE, comment=''}
new.df <- data.frame(capital=c(2.97, 10.45, 3.85), labour=c(85, 60, 105) )
stargazer(new.df, summary = FALSE, type = 'text', digit.separator = '')
```

Evaluate the prediction of the dependent variable for each individual.  
**Round the answer to `r my.digits` decimal places.**

The answer

```{r, echo=FALSE, comment=''}
my.predict <- exp( predict(regr, new.df) )
# names(my.predict) <- 1:nrow(new.df)
# round(my.predict, digits = my.digits )
stargazer(round(data.frame(Prediction=my.predict), digits = my.digits ), 
          type='text', summary=FALSE, digit.separator = '', digits=my.digits)
```

# output equation #2

For the dataset `Labour` consider a regression
**log(output) на log(capital), log(labour), log^2^(capital), log^2^(labour)**.

Fitting results:

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.digits.output <- 3
my.digits <- 2
regr <- lm(formula=log(output)~log(capital)+log(labour)+I(log(capital)^2)+I(log(labour)^2), data=Labour)
stargazer(regr, type='text', digits=my.digits.output, digit.separator = '', df=FALSE)
```

Consider firms with the following characteristics:

```{r, echo=FALSE, comment=''}
new.df <- data.frame(capital=c(22.14, 7.32, 0.67), labour=c(407, 197, 31) )
stargazer(new.df, summary = FALSE, type = 'text', digit.separator = '')
```

Evaluate the prediction of the dependent variable for each individual.  
**Round the answer to `r my.digits` decimal places.**

The answer

```{r, echo=FALSE, comment=''}
my.predict <- exp( predict(regr, new.df) )
# names(my.predict) <- 1:nrow(new.df)
# round(my.predict, digits = my.digits )
stargazer(round(data.frame(Prediction=my.predict), digits = my.digits ), 
          type='text', summary=FALSE, digit.separator = '', digits=my.digits)
```