---
title: "List 09: logit/probit-regression"
author: "N.V. Artamonov"
output:
  pdf_document:
    toc: yes
    number_sections: true
    df_print: default
    latex_engine: xelatex
  html_document:
    toc: true
    toc_float: 
        collapsed: false
    number_sections: true
    df_print: paged
header-includes:
   - \usepackage{xltxtra}
   - \usepackage{fontspec}
   - \setmainfont{Times New Roman}
   - \setsansfont{Arial}
   - \setmonofont{Courier New}
   - \newfontfamily{\cyrillicfont}{Times New Roman}
   - \newfontfamily{\cyrillicfonttt}{Courier New} 
   - \newfontfamily{\cyrillicfontsf}{Arial}
   - \usepackage{xcolor}
   - \usepackage[english,russian]{babel}
---

```{r, message=FALSE, echo=FALSE}
library(stargazer)
library(lmtest)
library(car)
library(sandwich)
library(jtools)
data(loanapp, package = 'wooldridge')
data(SwissLabor, package = 'AER')
mroz_Greene <- read.csv('https://raw.githubusercontent.com/artamonoff/econometrica/refs/heads/main/econometrica-2/datasets/TableF5-1.csv')
```

# Fitting & coefficients' interpretation

## approve equation #1 (probit)

For the dataset `loanapp` consider probit-regression **approve on appinc, mortno, unem, dep, male, married, yjob, self**

Specification: 
$P(approve=1)=\Phi(\beta_0+\beta_1appinc+\beta_2mortno+\beta_3unem+\beta_4dep+\beta_5male+\beta_6married+\beta_7yjob+\beta_8self)$

An alternative specification:
$probit(P(approve=1))=\beta_0+\beta_1appinc+\beta_2mortno+\beta_3unem+\beta_4dep+\beta_5male+\beta_6married+\beta_7yjob+\beta_8self$

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.digits <- 3
my.digits.output <- 3
# sign.level <- 0.10
regr <- glm(formula=approve~appinc+mortno+unem+dep+male+married+yjob+self, data=loanapp, 
            family = binomial(link='probit') )
```

Fit the model and report coefficinets. **Round the answer to `r my.digits` decimal places.**

The answer:

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
round( coef(regr), digits = 3)
```

Give the interpretation of coefficients.

## approve equation #2 (logit)

For the dataset `loanapp` consider logit-regression **approve on appinc, mortno, unem, dep, male, married, yjob, self**

Specification: 
$P(approve=1)=\Lambda(\beta_0+\beta_1appinc+\beta_2mortno+\beta_3unem+\beta_4dep+\beta_5male+\beta_6married+\beta_7yjob+\beta_8self)$

An alternative specification:
$logit(P(approve=1))=\beta_0+\beta_1appinc+\beta_2mortno+\beta_3unem+\beta_4dep+\beta_5male+\beta_6married+\beta_7yjob+\beta_8self$

Здесь \(logit(P(approve=1))=\log\frac{P(approve=1)}{1-P(approve=1)}=\log\frac{P(approve=1)}{P(approve=0)}\)

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.digits <- 3
my.digits.output <- 3
# sign.level <- 0.10
regr <- glm(formula=approve~appinc+mortno+unem+dep+male+married+yjob+self, data=loanapp, 
            family = binomial(link='logit') )
```

Fit the model and report coefficients. **Round the answer to `r my.digits` decimal places.**

The answer:

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
round( coef(regr), digits = my.digits)
```

Give the interpretation of coefficinets..

## labour force equation #1 (probit)

For the dataset `TableF5-1` consider probit-regression **LFP on WA, WA^2, WE, KL6, K618, CIT, UN, log(FAMINC)**

Specification: 
$P(LFP=1)=\Phi(\beta_0+\beta_1WA+\beta_2WA^2+\beta_3WE+\beta_4KL6+\beta_5K618+\beta_5CIT+\beta_7UN+\beta_8\log(FAMINC))$

An alternative specification:
$probit(P(LFP=1))=\beta_0+\beta_1WA+\beta_2WA^2+\beta_3WE+\beta_4KL6+\beta_5K618+\beta_5CIT+\beta_7UN+\beta_8\log(FAMINC)$

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.digits <- 3
my.digits.output <- 3
# sign.level <- 0.10
regr <- glm(formula=LFP~WA+I(WA^2)+WE+KL6+K618+CIT+UN+log(FAMINC), data=mroz_Greene, 
            family = binomial(link='probit') )
```

Fit the model and report coefficinets. **Round the answer to `r my.digits` decimal places.**

The answer:

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
round( coef(regr), digits = my.digits)
```

Give the interpretation of coefficients.

## labour force equation #2 (logit)

For the dataset `TableF5-1` consider logit-regression **LFP on WA, WA^2, WE, KL6, K618, CIT, UN, log(FAMINC)**

Specification: 
$P(LFP=1)=\Lambda(\beta_0+\beta_1WA+\beta_2WA^2+\beta_3WE+\beta_4KL6+\beta_5K618+\beta_5CIT+\beta_7UN+\beta_8\log(FAMINC))$

An alternative specification:
$logit(P(LFP=1))=\beta_0+\beta_1WA+\beta_2WA^2+\beta_3WE+\beta_4KL6+\beta_5K618+\beta_5CIT+\beta_7UN+\beta_8\log(FAMINC)$

Здесь \(logit(P(LFP=1))=\log\frac{P(LFP=1)}{1-P(LFP=1)}=\log\frac{P(LFP=1)}{P(LFP=0)}\)

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.digits <- 3
my.digits.output <- 3
# sign.level <- 0.10
regr <- glm(formula=LFP~WA+I(WA^2)+WE+KL6+K618+CIT+UN+log(FAMINC), data=mroz_Greene, 
            family = binomial(link='logit') )
```

Fit the model and report coefficients. **Round the answer to `r my.digits` decimal places.**

The answer:

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
round( coef(regr), digits = my.digits)
```

Give the interpretation of coefficinets..

# z-test

## approve equation #1 (probit)

For the dataset `loanapp` consider probit-regression **approve on appinc, mortno, unem, dep, male, married, yjob, self**

Fit the model and report the output result of z-test

The answer:

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.digits <- 3
my.digits.output <- 4
sign.level <- 0.10
regr <- glm(formula=approve~appinc+mortno+unem+dep+male+married+yjob+self, data=loanapp, 
            family = binomial(link='probit') )
round(coeftest(regr), digits = my.digits.output)
# z.test <- round(coeftest(regr), digits = my.digits.output)
# z.test[,1:2]
```

<span style="color: blue">Significance level is `r 100*sign.level`%</span>

Evaluate critical value. **Round the answer to `r my.digits` decimal places.**

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
z.cr <- qnorm(p=1-sign.level/2)
round(z.cr , digits = my.digits)
```

Which coefficiants are significant? The answer

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
z.test <- coeftest(regr)
regressors <- row.names(z.test)
sign.regressors <- NULL
for(i in 1:nrow(z.test) ) {
  if(z.test[i,4]<sign.level) sign.regressors <- c(sign.regressors, regressors[i])
}
print(sign.regressors)
```

## approve equation #2 (logit)

For the dataset `loanapp` consider logit-regression **approve on appinc, mortno, unem, dep, male, married, yjob, self**

Fit the model and report the output result of z-test

The answer:

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.digits <- 3
my.digits.output <- 4
sign.level <- 0.05
regr <- glm(formula=approve~appinc+mortno+unem+dep+male+married+yjob+self, data=loanapp, 
            family = binomial(link='logit') )
round(coeftest(regr), digits = my.digits.output)
```

 
<span style="color: blue">Significance level is `r 100*sign.level`%</span>

Evaluate critical value. **Round the answer to `r my.digits` decimal places.**

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
z.cr <- qnorm(p=1-sign.level/2)
round(z.cr , digits = my.digits)
```

Which coefficiants are significant? The answer

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
z.test <- coeftest(regr)
regressors <- row.names(z.test)
sign.regressors <- NULL
for(i in 1:nrow(z.test) ) {
  if(z.test[i,4]<sign.level) sign.regressors <- c(sign.regressors, regressors[i])
}
print(sign.regressors)
```

## labour force equation #1 (probit)

For the dataset `TableF5-1` consider probit-regression **LFP on WA, WA^2, WE, KL6, K618, CIT, UN, log(FAMINC)**

Fit the model and report the output result of z-test

The answer:

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.digits <- 3
my.digits.output <- 4
sign.level <- 0.10
regr <- glm(formula=LFP~WA+I(WA^2)+WE+KL6+K618+CIT+UN+log(FAMINC), data=mroz_Greene, 
            family = binomial(link='probit') )
round(coeftest(regr), digits = my.digits.output)
```

<span style="color: blue">Significance level is `r 100*sign.level`%</span>

Evaluate critical value. **Round the answer to `r my.digits` decimal places.**

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
z.cr <- qnorm(p=1-sign.level/2)
round(z.cr , digits = my.digits)
```

Which coefficiants are significant? The answer

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
z.test <- coeftest(regr)
regressors <- row.names(z.test)
sign.regressors <- NULL
for(i in 1:nrow(z.test) ) {
  if(z.test[i,4]<sign.level) sign.regressors <- c(sign.regressors, regressors[i])
}
print(sign.regressors)
```

## labour force equation #2 (logit)

For the dataset `TableF5-1` consider logit-regression **LFP on WA, WA^2, WE, KL6, K618, CIT, UN, log(FAMINC)**

Fit the model and report the output result of z-test

The answer:

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.digits <- 3
my.digits.output <- 4
sign.level <- 0.05
regr <- glm(formula=LFP~WA+I(WA^2)+WE+KL6+K618+CIT+UN+log(FAMINC), data=mroz_Greene, 
            family = binomial(link='logit') )
round(coeftest(regr), digits = my.digits.output)
```

<span style="color: blue">Significance level is `r 100*sign.level`%</span>

Evaluate critical value. **Round the answer to `r my.digits` decimal places.**

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
z.cr <- qnorm(p=1-sign.level/2)
round(z.cr , digits = my.digits)
```

Which coefficiants are significant? The answer

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
z.test <- coeftest(regr)
regressors <- row.names(z.test)
sign.regressors <- NULL
for(i in 1:nrow(z.test) ) {
  if(z.test[i,4]<sign.level) sign.regressors <- c(sign.regressors, regressors[i])
}
print(sign.regressors)
```

# LR-test: overall significance

## approve equation #1 (probit)

For the dataset `loanapp` consider probit-regression **approve on appinc, unem, male, yjob, self**

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.digits <- 3
my.digits.output <- 4
sign.level <- 0.10
regr <- glm(formula=approve~appinc+unem+male+yjob+self, data=loanapp, 
            family = binomial(link='probit') )
# summ(regr, digits = my.digits.output)
my.logL <- round(-deviance(regr)/2, digits = my.digits.output)
my.logL.null <- round(-regr$null.deviance/2, digits = my.digits.output)
```

Fit the model and test overall significance.
<span style="color: blue">Significance level is `r 100*sign.level`%</span>.

Evaluate test statistics. **Round the answer to `r my.digits` decimal places.**

The answer:

```{r, echo=FALSE, comment=''}
LR.stat <- regr$null.deviance-regr$deviance
round(LR.stat, digits = my.digits)
```

Evaluate critical value. **Round the answer to `r my.digits` decimal places.**

```{r, echo=FALSE, comment=''}
chi.sq.cr <- qchisq(p=1-sign.level, df=regr$rank-1)
round(chi.sq.cr, digits = my.digits)
```

Is the regression significant? The answer

```{r, echo=FALSE, comment=''}
if(LR.stat>chi.sq.cr) print('Significant') else print('Insignificant')
```

Which coefficinets are significante?

```{r, echo=FALSE, comment=''}
z.test <- coeftest(regr)
regressors <- row.names(z.test)
sign.regressors <- NULL
for(i in 1:nrow(z.test) ) {
  if(z.test[i,4]<sign.level) sign.regressors <- c(sign.regressors, regressors[i])
}
print(sign.regressors)
```

## approve equation #2 (logit)

For the dataset `loanapp` consider logit-regression **approve on appinc, appinc^2, mortno, unem, dep, male, married, yjob, self**

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.digits <- 3
my.digits.output <- 4
sign.level <- 0.05
regr <- glm(formula=approve~appinc+I(appinc^2)+mortno+unem+dep+male+married+yjob+self, data=loanapp, 
            family = binomial(link='logit') )
# summ(regr, digits = my.digits.output)
my.logL <- round(-deviance(regr)/2, digits = my.digits.output)
my.logL.null <- round(-regr$null.deviance/2, digits = my.digits.output)
```

Fit the model and test overall significance.
<span style="color: blue">Significance level is `r 100*sign.level`%</span>.

Evaluate test statistics. **Round the answer to `r my.digits` decimal places.**

The answer:

```{r, echo=FALSE, comment=''}
LR.stat <- regr$null.deviance-regr$deviance
round(LR.stat, digits = my.digits)
```

Evaluate critical value. **Round the answer to `r my.digits` decimal places.**

```{r, echo=FALSE, comment=''}
chi.sq.cr <- qchisq(p=1-sign.level, df=regr$rank-1)
round(chi.sq.cr, digits = my.digits)
```

Is regression significant? The answer

```{r, echo=FALSE, comment=''}
if(LR.stat>chi.sq.cr) print('Significant') else print('Insignificant')
```

Which coefficinets are significante?

```{r, echo=FALSE, comment=''}
z.test <- coeftest(regr)
regressors <- row.names(z.test)
sign.regressors <- NULL
for(i in 1:nrow(z.test) ) {
  if(z.test[i,4]<sign.level) sign.regressors <- c(sign.regressors, regressors[i])
}
print(sign.regressors)
```

## labour force equation #1 (probit)

For the dataset `TableF5-1` consider a collection of probit-regression. Fitting results

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.digits <- 3
my.digits.output <- 4
sign.level <- 0.10
regr <- NULL
regr[[1]] <- glm(formula=LFP~WA+I(WA^2)+WE+KL6+K618+CIT+UN+log(FAMINC), data=mroz_Greene,
                 family = binomial(link='probit') )
regr[[2]] <- update(regr[[1]], formula. = ~.-WE-KL6-K618)
regr[[3]] <- update(regr[[1]], formula. = ~.-WE-KL6-K618-log(FAMINC))
regr[[4]] <- update(regr[[1]], formula. = ~.-WE-KL6-K618-log(FAMINC)-WA-I(WA^2))
stargazer(regr, type='text', digits=my.digits.output, digit.separator = '', 
          dep.var.caption = 'Dependent variable', df=FALSE)
my.logL.null <- -regr[[1]]$null.deviance/2
my.logL <- NULL
for(i in 1:length(regr)) my.logL <- c(my.logL, -deviance(regr[[i]])/2 )
```

For each regression evauate LR-statistics for overall significance.
**Round the answer to `r my.digits` decimal places.**

The answer

```{r, echo=FALSE, comment=''}
LR.stat <- NULL
for(i in 1:length(regr)) {
  LR.stat <- c(LR.stat, 2*(my.logL[[i]]-my.logL.null) )
}
df.LR <- data.frame(Regression=1:length(regr), LR.stat= LR.stat) # round(LR.stat, digits = my.digits)
stargazer(df.LR, summary = FALSE, type='text', rownames=FALSE, digits = my.digits)
```

For each regression evaluate necessary critical value.
<span style="color: blue">Significance level is `r 100*sign.level`%</span>.
**Round the answer to `r my.digits` decimal places.**

The answer

```{r, echo=FALSE, comment=''}
chi.sq.cr <- NULL
for(i in 1:length(regr)) {
  chi.sq.cr <- c(chi.sq.cr, qchisq(p=1-sign.level, df=regr[[i]]$rank-1) )
}
df.chi.sq <- data.frame(Regression=1:length(regr), Critical= chi.sq.cr) # round(LR.stat, digits = my.digits)
stargazer(df.chi.sq, summary = FALSE, type='text', rownames=FALSE, digits = my.digits)
```

Which regression is significant?

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
significance <- NULL
for(i in 1:length(regr)) {
  if( LR.stat[i]>chi.sq.cr[i]) significance <- c(significance, 'Significant')
  else significance <- c(significance, 'Insignificant')
}
df.sign <- data.frame(Regression=1:length(regr), Significance=significance)
stargazer(df.sign, summary = FALSE, type='text', rownames=FALSE)
```

## labour force equation #2 (logit)

For the dataset `TableF5-1` consider a collection of logit-regressions. fitting results

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.digits <- 3
my.digits.output <- 4
sign.level <- 0.05
regr <- NULL
regr[[1]] <- glm(formula=LFP~WA+I(WA^2)+WE+KL6+K618+CIT+UN+log(FAMINC), data=mroz_Greene,
                 family = binomial(link='logit') )
regr[[2]] <- update(regr[[1]], formula. = ~.-WA-I(WA^2) )
regr[[3]] <- update(regr[[1]], formula. = ~.-WA-I(WA^2)-WE )
regr[[4]] <- update(regr[[1]], formula. = ~.-WA-I(WA^2)-WE-log(FAMINC)-KL6)
stargazer(regr, type='text', digits=my.digits.output, digit.separator = '', 
          dep.var.caption = 'Dependent variable', df=TRUE )
my.logL.null <- -regr[[1]]$null.deviance/2
my.logL <- NULL
for(i in 1:length(regr)) my.logL <- c(my.logL, -deviance(regr[[i]])/2 )
```

For each regression evauate LR-statistics for overall significance.
**Round the answer to `r my.digits` decimal places.**

The answer

```{r, echo=FALSE, comment=''}
LR.stat <- NULL
for(i in 1:length(regr)) {
  LR.stat <- c(LR.stat, 2*(my.logL[[i]]-my.logL.null) )
}
df.LR <- data.frame(Regression=1:length(regr), LR.stat= LR.stat) # round(LR.stat, digits = my.digits)
stargazer(df.LR, summary = FALSE, type='text', rownames=FALSE, digits = my.digits)
```

For each regression evaluate necessary critical value.
<span style="color: blue">Significance level is `r 100*sign.level`%</span>.
**Round the answer to `r my.digits` decimal places.**

The answer

```{r, echo=FALSE, comment=''}
chi.sq.cr <- NULL
for(i in 1:length(regr)) {
  chi.sq.cr <- c(chi.sq.cr, qchisq(p=1-sign.level, df=regr[[i]]$rank-1) )
}
df.chi.sq <- data.frame(Regression=1:length(regr), Critical= chi.sq.cr) # round(LR.stat, digits = my.digits)
stargazer(df.chi.sq, summary = FALSE, type='text', rownames=FALSE, digits = my.digits)
```

Which regression is significant?

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
significance <- NULL
for(i in 1:length(regr)) {
  if( LR.stat[i]>chi.sq.cr[i]) significance <- c(significance, 'Significant')
  else significance <- c(significance, 'Insignificant')
}
df.sign <- data.frame(Regression=1:length(regr), Significance=significance)
stargazer(df.sign, summary = FALSE, type='text', rownames=FALSE )
```

# Wald-test: joint significance

## swiss labour force equation #1

For the dataset `SwissLabour` consider logit-regression **participation on income, income^2, age, age^2, youngkids, oldkids, foreign**

Fitting result

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.digits <- 3
my.digits.output <- 4
sign.level <- 0.05
regr <- glm(formula=as.numeric(participation)-1~income+I(income^2)+age+I(age^2)+youngkids+oldkids+foreign, 
            data=SwissLabor, 
            family = binomial(link='logit') )
summ(regr, digits = my.digits.output, vifs=TRUE)
```

Fit the model and test the following hypothesis with Wlad test (use \(\chi^2\)-statistics).
<span style="color: blue">Significance level is `r 100*sign.level`%</span>.

### Hypothesis 1

Test the significance of income, i.e. test the hypothesis \(H_0:\beta_{income}=\beta_{income^2}=0\).

Evaluate test statistics and its P-value.

```{r,echo=FALSE, comment=''}
regr.rest <- update(regr, formula. = ~.-income-I(income^2) )
w.test <- waldtest(regr.rest, regr, test='Chisq')
stargazer(w.test[2,3:4], type = 'text', summary = FALSE, rownames = FALSE)
```

Evaluate critical value. **Round the answer to `r my.digits` decimal places.**

```{r, echo=FALSE, comment=''}
ch2.cr <- qchisq(p=1-sign.level, df=w.test[2,2])
round(ch2.cr, digits = my.digits)
```

Conclusion

```{r, echo=FALSE, comment=''}
if(w.test[2,4]<sign.level) print('Significant') else print('Insignificant')
```

### Hypothesis 2

Test the significence of the numter of kids, i.e. the hypothesis \(H_0:\beta_{youngkids}=\beta_{oldkids}=0\).

Evaluate test statistics and its P-value.

```{r,echo=FALSE, comment=''}
regr.rest <- update(regr, formula. = ~.-youngkids-oldkids )
w.test <- waldtest(regr.rest, regr, test='Chisq')
stargazer(w.test[2,3:4], type = 'text', summary = FALSE, rownames = FALSE)
```

Evaluate critical value. **Round the answer to `r my.digits` decimal places.**

```{r, echo=FALSE, comment=''}
ch2.cr <- qchisq(p=1-sign.level, df=w.test[2,2])
round(ch2.cr, digits = my.digits)
```

Conclusion

```{r, echo=FALSE, comment=''}
if(w.test[2,4]<sign.level) print('Significant') else print('Insignificant')
```

### Hypothesis 3

Test the significence of age, i.e. the hypothesis \(H_0:\beta_{age}=\beta_{age^2}=0\).

Evaluate test statistics and its P-value.

```{r,echo=FALSE, comment=''}
regr.rest <- update(regr, formula. = ~.-age-I(age^2) )
w.test <- waldtest(regr.rest, regr, test='Chisq')
stargazer(w.test[2,3:4], type = 'text', summary = FALSE, rownames = FALSE)
```

Evaluate critical value. **Round the answer to `r my.digits` decimal places.**

```{r, echo=FALSE, comment=''}
ch2.cr <- qchisq(p=1-sign.level, df=w.test[2,2])
round(ch2.cr, digits = my.digits)
```

Conclusion

```{r, echo=FALSE, comment=''}
if(w.test[2,4]<sign.level) print('Significant') else print('Insignificant')
```