---
title: 'List 06: Multicollinearity'
author: "Nikita V. Artamonov"
output:
  pdf_document:
    toc: yes
    df_print: default
    latex_engine: xelatex
  html_document:
    toc: yes
    toc_float:
      collapsed: no
    number_sections: yes
    df_print: paged
---

```{r, message=FALSE, echo=FALSE}
library(stargazer)
library(lmtest)
library(car)
library(ggplot2)
library(corrplot)
library(ggcorrplot)
library(sandwich)
data(sleep75, wage1, wage2, Labour, Electricity, package = c('wooldridge', 'Ecdat') )
sleep75$gender <- factor(sleep75$male, labels = c('female', 'male') )
wage1$gender <- factor(wage1$female, labels = c('male', 'female') )
```

# sleep equation #1

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.digits.output <- 3
my.digits <- 2
sign.level <- 0.05

df.name <- 'sleep75'
df <- get(df.name)
spec <- formula('sleep~1+I(totwrk/100)+I(totwrk^2/10000)+age+smsa+male')

mod <- lm(formula=spec, data=df)
```

For the dataset ` `r df.name` ` consider a regreaion **`r format(spec)`**.

Evaluate VIF for each regressors

```{r, echo=FALSE, comment=''}
vif(mod)
```

Calculate correlation matirx for regressosrs

```{r, echo=FALSE, comment=''}
round(cor(model.frame(mod)[,-1]), digits = my.digits.output )
```

and visualize it

```{r, echo=FALSE, comment=''}
ggcorrplot(cor(model.frame(mod)[,-1]), method = "square", lab=TRUE)
```

<span style="color: blue">Significant level is `r 100*sign.level`%</span>. Which coefficients are sognoficant 
(perform non-robust t-test)?

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
t.test <- coeftest(mod)
regressors <- row.names(t.test)
sign.regressors <- NULL
for(i in 2:nrow(t.test) ) {
  if(t.test[i,4]<sign.level) sign.regressors <- c(sign.regressors, regressors[i])
}
print(sign.regressors)
```

We test the significance of working time, i.e. the hypothesis \(H_0:\beta_{totwrk/100}=\beta_{totwrk^2/10000}=0\).
Testing result (Non-robust test):

```{r,comment='', echo=FALSE}
testing.result <- linearHypothesis(mod, c('I(totwrk/100)=0', 'I(totwrk^2/10000)=0'))
# testing.result
stargazer(testing.result[2,5:6], summary = FALSE, type = 'text', rownames=FALSE)
```

Calculate the required critical value. **Round to `r my.digits` decimal places.**

```{r, echo=FALSE, comment=''}
F.cr <- qf(p=1-sign.level, df1 = testing.result[2,3], df2 = df.residual(mod))
round(F.cr, digits = my.digits)
```

Inferences:

```{r,comment='', echo=FALSE}
if(testing.result[2,6]<sign.level) print('We reject the null hypothesis') else print('We do not reject the null hypothesis')
```

**At first glance we have a contradiction.** It is caused by multicollinearity.

# sleep equation #2

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.digits.output <- 3
my.digits <- 2
sign.level <- 0.05

df.name <- 'sleep75'
df <- get(df.name)
spec <- formula('sleep~totwrk+age+I(age^2)+smsa+male+union')

mod <- lm(formula=spec, data=df)
```

For the dataset ` `r df.name` ` consider a regreaion **`r format(spec)`**.

Evaluate VIF for each regressors

```{r, echo=FALSE, comment=''}
vif(mod)
```

Calculate correlation matirx for regressosrs

```{r, echo=FALSE, comment=''}
round(cor(model.frame(mod)[,-1]), digits = my.digits.output )
```

and visualize it

```{r, echo=FALSE, comment=''}
ggcorrplot(cor(model.frame(mod)[,-1]), method = "square", lab=TRUE)
```

<span style="color: blue">Significant level is `r 100*sign.level`%</span>. Which coefficients are sognoficant 
(perform non-robust t-test)?

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
t.test <- coeftest(mod)
regressors <- row.names(t.test)
sign.regressors <- NULL
for(i in 2:nrow(t.test) ) {
  if(t.test[i,4]<sign.level) sign.regressors <- c(sign.regressors, regressors[i])
}
print(sign.regressors)
```

We test the significance of age, i.e. the hypothesis \(H_0:\beta_{age}=\beta_{age^2}=0\).
Testing result (Non-robust test):

```{r,comment='', echo=FALSE}
testing.result <- linearHypothesis(mod, c('age=0', 'I(age^2)=0'))
# testing.result
stargazer(testing.result[2,5:6], summary = FALSE, type = 'text', rownames=FALSE)
```

Calculate the required critical value. **Round to `r my.digits` decimal places.**

```{r, echo=FALSE, comment=''}
F.cr <- qf(p=1-sign.level, df1 = testing.result[2,3], df2 = df.residual(mod))
round(F.cr, digits = my.digits)
```

Inferences:

```{r,comment='', echo=FALSE}
if(testing.result[2,6]<sign.level) print('We reject the null hypothesis') else print('We do not reject the null hypothesis')
```

**At first glance we have a contradiction.** It is caused by multicollinearity.

# sleep equation #3

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.digits.output <- 3
my.digits <- 2
sign.level <- 0.05

df.name <- 'sleep75'
df <- get(df.name)
# spec <- formula('sleep~totwrk+south+I(totwrk*south)+age+male+smsa+marr')
spec <- formula('sleep~totwrk+age+smsa+south+I(totwrk*south)+I(age*south)+I(smsa*south)')

mod <- lm(formula=spec, data=df)
```

For the dataset ` `r df.name` ` consider a regreaion **`r format(spec)`**.

Evaluate VIF for each regressors

```{r, echo=FALSE, comment=''}
vif(mod)
```

Calculate correlation matirx for regressosrs

```{r, echo=FALSE, comment=''}
round(cor(model.frame(mod)[,-1]), digits = my.digits.output )
```

and visualize it

```{r, echo=FALSE, comment=''}
ggcorrplot(cor(model.frame(mod)[,-1]), method = "square", lab=TRUE)
```

<span style="color: blue">Significant level is `r 100*sign.level`%</span>. Which coefficients are sognoficant 
(perform non-robust t-test)?

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
t.test <- coeftest(mod)
regressors <- row.names(t.test)
sign.regressors <- NULL
for(i in 2:nrow(t.test) ) {
  if(t.test[i,4]<sign.level) sign.regressors <- c(sign.regressors, regressors[i])
}
print(sign.regressors)
```

We test the significance of geographical dummy, i.e. the hypothesis 
\(H_0:\beta_{south}=\beta_{totwrk*south}=\beta_{age * south}=\beta_{smsa * south}=0\).
Testing result (Non-robust test):

```{r,comment='', echo=FALSE}
testing.result <- linearHypothesis(mod, c('south=0', 'I(totwrk * south)=0', 'I(age * south)=0', 'I(smsa * south)=0'))
# testing.result
stargazer(testing.result[2,5:6], summary = FALSE, type = 'text', rownames=FALSE)
```

Calculate the required critical value. **Round to `r my.digits` decimal places.**

```{r, echo=FALSE, comment=''}
F.cr <- qf(p=1-sign.level, df1 = testing.result[2,3], df2 = df.residual(mod))
round(F.cr, digits = my.digits)
```

Inferences:

```{r,comment='', echo=FALSE}
if(testing.result[2,6]<sign.level) print('We reject the null hypothesis') else print('We do not reject the null hypothesis')
```

**At first glance we have a contradiction.** It is caused by multicollinearity.

# wage equation #1

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.digits.output <- 3
my.digits <- 2
sign.level <- 0.05

df.name <- 'wage2'
df <- get(df.name)
spec <- formula('log(wage)~age+I(age^2)+IQ+married+south+urban')

mod <- lm(formula=spec, data=df)
```

For the dataset ` `r df.name` ` consider a regreaion **`r format(spec)`**.

Evaluate VIF for each regressors

```{r, echo=FALSE, comment=''}
vif(mod)
```

Calculate correlation matirx for regressosrs

```{r, echo=FALSE, comment=''}
round(cor(model.frame(mod)[,-1]), digits = my.digits.output )
```

and visualize it

```{r, echo=FALSE, comment=''}
ggcorrplot(cor(model.frame(mod)[,-1]), method = "square", lab=TRUE)
```

<span style="color: blue">Significant level is `r 100*sign.level`%</span>. Which coefficients are sognoficant 
(perform non-robust t-test)?

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
t.test <- coeftest(mod)
regressors <- row.names(t.test)
sign.regressors <- NULL
for(i in 2:nrow(t.test) ) {
  if(t.test[i,4]<sign.level) sign.regressors <- c(sign.regressors, regressors[i])
}
print(sign.regressors)
```

We test the significance of age, i.e. the hypothesis \(_0:\beta_{age}=\beta_{age^2}=0\).
Testing result (Non-robust test):

```{r,comment='', echo=FALSE}
testing.result <- linearHypothesis(mod, c('age=0', 'I(age^2)=0'))
# testing.result
stargazer(testing.result[2,5:6], summary = FALSE, type = 'text', rownames=FALSE)
```

Calculate the required critical value. **Round to `r my.digits` decimal places.**

```{r, echo=FALSE, comment=''}
F.cr <- qf(p=1-sign.level, df1 = testing.result[2,3], df2 = df.residual(mod))
round(F.cr, digits = my.digits)
```

Inferences:

```{r,comment='', echo=FALSE}
if(testing.result[2,6]<sign.level) print('We reject the null hypothesis') else print('We do not reject the null hypothesis')
```

**At first glance we have a contradiction.** It is caused by multicollinearity.

# wage equation #2 (structural breaks)

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.digits.output <- 3
my.digits <- 2
sign.level <- 0.01

df.name <- 'wage2'
df <- get(df.name)
spec <- formula('log(wage)~age+IQ+south+urban+I(age*urban)+I(IQ*urban)+I(south*urban)')

mod <- lm(formula=spec, data=df)
```

Для набора данных ` `r df.name` ` рассмотрим линейную регрессию
**`r format(spec)`**.

For the dataset ` `r df.name` ` consider a regreaion **`r format(spec)`**.

Evaluate VIF for each regressors

```{r, echo=FALSE, comment=''}
vif(mod)
```

Calculate correlation matirx for regressosrs

```{r, echo=FALSE, comment=''}
round(cor(model.frame(mod)[,-1]), digits = my.digits.output )
```

and visualize it

```{r, echo=FALSE, comment=''}
ggcorrplot(cor(model.frame(mod)[,-1]), method = "square", lab=TRUE)
```

<span style="color: blue">Significant level is `r 100*sign.level`%</span>. Which coefficients are sognoficant 
(perform non-robust t-test)?

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
t.test <- coeftest(mod)
regressors <- row.names(t.test)
sign.regressors <- NULL
for(i in 2:nrow(t.test) ) {
  if(t.test[i,4]<sign.level) sign.regressors <- c(sign.regressors, regressors[i])
}
print(sign.regressors)
```

We test the significance of dwelling dummy, i.e. the hypothesis
\(H_0:\beta_{urban}=\beta_{age*urban}=\beta_{IQ*urban}=\beta_{south*urban}=0\).
Testing result (Non-robust test):

```{r,comment='', echo=FALSE}
testing.result <- linearHypothesis(mod, c('urban=0', 'I(age * urban)=0', 'I(IQ * urban)=0', 'I(south * urban)=0'))
# testing.result
stargazer(testing.result[2,5:6], summary = FALSE, type = 'text', rownames=FALSE)
```

Calculate the required critical value. **Round to `r my.digits` decimal places.**

```{r, echo=FALSE, comment=''}
F.cr <- qf(p=1-sign.level, df1 = testing.result[2,3], df2 = df.residual(mod))
round(F.cr, digits = my.digits)
```

Inferences:

```{r,comment='', echo=FALSE}
if(testing.result[2,6]<sign.level) print('We reject the null hypothesis') else print('We do not reject the null hypothesis')
```

**At first glance we have a contradiction.** It is caused by multicollinearity.