---
title: "List 11: Marginal effects"
author: "N.V. Artamonov"
output:
  pdf_document:
    toc: yes
    number_sections: true
    df_print: default
    latex_engine: xelatex
  html_document:
    toc: true
    toc_float: 
        collapsed: false
    number_sections: true
    df_print: paged
header-includes:
   - \usepackage{xltxtra}
   - \usepackage{fontspec}
   - \setmainfont{Times New Roman}
   - \setsansfont{Arial}
   - \setmonofont{Courier New}
   - \newfontfamily{\cyrillicfont}{Times New Roman}
   - \newfontfamily{\cyrillicfonttt}{Courier New} 
   - \newfontfamily{\cyrillicfontsf}{Arial}
   - \usepackage{xcolor}
   - \usepackage[english,russian]{babel}
---

```{r, message=FALSE, echo=FALSE}
library(stargazer)
library(lmtest)
library(car)
library(sandwich)
library(jtools)
library(DescTools)
library(erer)
data(loanapp, SwissLabor, Mroz, package=c('wooldridge', 'AER', 'Ecdat'))
# data(loanapp, package = 'wooldridge')
# data(SwissLabor, package = 'AER')
mroz_Greene <- read.csv('https://raw.githubusercontent.com/artamonoff/econometrica/refs/heads/main/econometrica-2/datasets/TableF5-1.csv')
```

# Marginal effects 

## for probit

Consider a probit-regression $P(y=1)=\Phi(x'\beta)$, where
$x'\beta=\beta_0+\beta_1x_1+\cdots+\beta_kx_k$

Marginal effects $\frac{\partial P(y=1)}{\partial x_j}=\phi(x'\beta)\beta_j$,
where $\phi(z)=\frac{1}{\sqrt{2\pi}}exp\{-z^2/2\}$

## for logit

Consider a logit-regression $P(y=1)=\Lambda(x'\beta)$, where
$x'\beta=\beta_0+\beta_1x_1+\cdots+\beta_kx_k$

Marginal effects $\frac{\partial P(y=1)}{\partial x_j}=\lambda(x'\beta)\beta_j$,
where $\lambda(z)=\frac{exp(z)}{(1+exp(z))^2}$

## Average marginal effects

Usually we consider 

- marginal effects at the mean of each regressor : 
  * \(\phi(\bar{x}'\beta)\beta_j\) for probit 
  * \(\lambda(\bar{x}'\beta)\beta_j\) for logit
- The average of the marginal effects at each observation: 
  * \(\overline{\phi(x'\beta)\beta_j}\)  for probit 
  * \(\overline{\lambda(x'\beta)\beta_j}\) for logit

# labour force equation

For the dataset `TableF5-1` consider regression  **LFP на WA, log(FAMINC), WE, KL6, K618, CIT, UN**
of the following specifications:

- logit
- probit

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.digits <- 3
my.digits.output <- 3
spec <- formula(LFP~WA+log(FAMINC)+WE+KL6+K618+CIT+UN)
my.df <- mroz_Greene
# regr.LPM <- lm(formula=spec, data=my.df)
regr.logit <- glm(formula=spec, data=my.df, family=binomial(link='logit'), x=TRUE)
regr.probit <- glm(formula=spec, data=my.df, family=binomial(link='probit'), x=TRUE)
rm(my.df)
# stargazer(regr.LPM, regr.logit, regr.probit, type='text', digits=my.digits.output, digit.separator = '', 
#           dep.var.caption = 'Зависимая переменная', df=FALSE)
```

The marginal effects at the mean of each regressor for logit model 
**Round the answer to `r my.digits` decimal places.**

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
maBina(regr.logit, x.mean=TRUE, digits = my.digits)
```

The marginal effects at the mean of each regressor for probit model. 
**Round the answer to `r my.digits` decimal places.**

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
maBina(regr.probit, x.mean=TRUE, digits = my.digits)
```

The average of the marginal effects at each observation for logit модели. 
**Round the answer to `r my.digits` decimal places.**

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
maBina(regr.logit, x.mean=FALSE, digits = my.digits)
```

The average of the marginal effects at each observation for probit модели. 
**Round the answer to `r my.digits` decimal places.**

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
maBina(regr.probit, x.mean=FALSE, digits = my.digits)
```
       
# approve equation

For the dataset `loanapp` consider regression **approve на appinc/100, mortno, unem, dep, male**
of the following specifications:

- logit
- probit

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.digits <- 3
my.digits.output <- 3
spec <- formula(approve~I(appinc/100)+mortno+unem+dep+male)
my.df <- loanapp
# regr.LPM <- lm(formula=spec, data=my.df)
regr.logit <- glm(formula=spec, data=my.df, family=binomial(link='logit'), x=TRUE)
regr.probit <- glm(formula=spec, data=my.df, family=binomial(link='probit'), x=TRUE)
rm(my.df)
# stargazer(regr.LPM, regr.logit, regr.probit, type='text', digits=my.digits.output, digit.separator = '', 
#           dep.var.caption = 'Зависимая переменная', df=FALSE)
```

The marginal effects at the mean of each regressor for logit model 
**Round the answer to `r my.digits` decimal places.**

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
maBina(regr.logit, x.mean=TRUE, digits = my.digits)
```

The marginal effects at the mean of each regressor for probit model. 
**Round the answer to `r my.digits` decimal places.**

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
maBina(regr.probit, x.mean=TRUE, digits = my.digits)
```

The average of the marginal effects at each observation for logit модели. 
**Round the answer to `r my.digits` decimal places.**

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
maBina(regr.logit, x.mean=FALSE, digits = my.digits)
```

The average of the marginal effects at each observation for probit модели. 
**Round the answer to `r my.digits` decimal places.**

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
maBina(regr.probit, x.mean=FALSE, digits = my.digits)
```

# swiss labour force equation

For the dataset `SwissLabour` consider regression **participation на income, age, age^2, youngkids, oldkids**
of the following specification:

- logit
- probit

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.digits <- 3
my.digits.output <- 3
spec <- formula(as.numeric(participation)-1~income+age+I(age^2)+youngkids+oldkids)
my.df <- SwissLabor
# regr.LPM <- lm(formula=spec, data=my.df)
regr.logit <- glm(formula=spec, data=my.df, family=binomial(link='logit'), x=TRUE)
regr.probit <- glm(formula=spec, data=my.df, family=binomial(link='probit'), x=TRUE)
rm(my.df)
# stargazer(regr.LPM, regr.logit, regr.probit, type='text', digits=my.digits.output, digit.separator = '', 
#           dep.var.caption = 'Зависимая переменная', df=FALSE)
```

The marginal effects at the mean of each regressor for logit model 
**Round the answer to `r my.digits` decimal places.**

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
maBina(regr.logit, x.mean=TRUE, digits = my.digits)
```

The marginal effects at the mean of each regressor for probit model. 
**Round the answer to `r my.digits` decimal places.**

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
maBina(regr.probit, x.mean=TRUE, digits = my.digits)
```

The average of the marginal effects at each observation for logit модели. 
**Round the answer to `r my.digits` decimal places.**

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
maBina(regr.logit, x.mean=FALSE, digits = my.digits)
```

The average of the marginal effects at each observation for probit модели. 
**Round the answer to `r my.digits` decimal places.**

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
maBina(regr.probit, x.mean=FALSE, digits = my.digits)