---
title: 'List 07: Heteroskedasticity'
author: "Nikita V. Artamonov"
output:
  pdf_document:
    toc: yes
    df_print: default
    latex_engine: xelatex
  html_document:
    toc: yes
    number_sections: yes
    df_print: paged
  word_document:
    toc: yes
---

```{r, message=FALSE, echo=FALSE}
library(stargazer)
library(lmtest)
library(car)
library(ggplot2)
library(sandwich)
data(sleep75, wage1, wage2, Labour, Electricity, Diamond, diamonds, package = c('wooldridge', 'Ecdat', 'ggplot2') )
sleep75$gender <- factor(sleep75$male, labels = c('female', 'male') )
wage1$gender <- factor(wage1$female, labels = c('male', 'female') )
```

# Tests

## sleep equation

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.digits.output <- 3
my.digits <- 2
sign.level <- 0.05
df.name <- 'sleep75'
df <- get(df.name)
spec <- formula('sleep~totwrk+age+I(age^2)+male+smsa+south')

mod <- lm(formula=spec, data=df)
```

For the dataset ` `r df.name` ` consider a regression
**`r format(spec)`**. Its specification is
\(sleep=\beta_0+\beta_1totwrk+\beta_2age+\beta_3age^2+\beta_4male+\beta_5smsa+\beta_6south+u\).

BP-test is based on the following heteroskedasicity model \(Var(u|X)=f(x'\gamma)\)

We test \(H_0:\gamma_1=\cdots=\gamma_k=0\) (homoskedastic error term)

Fit the regression and perform BP-test. Result is

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.bp.test <- bptest(mod)
my.bp.test
```

<span style="color: blue">significant level `r 100*sign.level`%</span>.

Evaluate necessary critical value  **Round the answer to `r my.digits` decimal places.**

```{r, echo=FALSE, comment=''}
chi2.cr <- qchisq(p=1-sign.level, df=my.bp.test$parameter[[1]] )
round(chi2.cr, digits = my.digits)
```

Conclusion

```{r echo=FALSE, comment=''}
if(my.bp.test$p.value[[1]]<sign.level) print('Evidence for heteroskedasticity') else print('Evidence for homoskedasticity')
```

## wage equation

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.digits.output <- 3
my.digits <- 2
sign.level <- 0.05
df.name <- 'wage1'
df <- get(df.name)
spec <- formula('log(wage)~exper+I(exper^2)+female+married+smsa')

mod <- lm(formula=spec, data=df)
```

For the dataset ` `r df.name` ` consider a regression
**`r format(spec)`**.

BP-test is based on the following heteroskedasicity model \(Var(u|X)=f(x'\gamma)\)

We test \(H_0:\gamma_1=\cdots=\gamma_k=0\) (homoskedastic error term)

Fit the regression and perform BP-test. Result is

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.bp.test <- bptest(mod)
my.bp.test
```

<span style="color: blue">significant level `r 100*sign.level`%</span>.

Evaluate necessary critical value  **Round the answer to `r my.digits` decimal places.**

```{r, echo=FALSE, comment=''}
chi2.cr <- qchisq(p=1-sign.level, df=my.bp.test$parameter[[1]] )
round(chi2.cr, digits = my.digits)
```

Conclusion

```{r echo=FALSE, comment=''}
if(my.bp.test$p.value[[1]]<sign.level) print('Evidence for heteroskedasticity') else print('Evidence for homoskedasticity')
```

## output equation

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.digits.output <- 3
my.digits <- 2
sign.level <- 0.05
df.name <- 'Labour'
df <- get(df.name)
spec <- formula('log(output)~log(capital)+log(labour)+I(log(capital)^2)+I(log(labour)^2)')

mod <- lm(formula=spec, data=df)
```

For the dataset ` `r df.name` ` consider a regression
**`r format(spec)`**. 

BP-test is based on the following heteroskedasicity model \(Var(u|X)=f(x'\gamma)\)

We test \(H_0:\gamma_1=\cdots=\gamma_k=0\) (homoskedastic error term)

Fit the regression and perform BP-test. Result is

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.bp.test <- bptest(mod)
my.bp.test
```

<span style="color: blue">significant level `r 100*sign.level`%</span>.

Evaluate necessary critical value  **Round the answer to `r my.digits` decimal places.**

```{r, echo=FALSE, comment=''}
chi2.cr <- qchisq(p=1-sign.level, df=my.bp.test$parameter[[1]] )
round(chi2.cr, digits = my.digits)
```

Conclusion

```{r echo=FALSE, comment=''}
if(my.bp.test$p.value[[1]]<sign.level) print('Evidence for heteroskedasticity') else print('Evidence for homoskedasticity')
```

## cost equation #1

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.digits.output <- 3
my.digits <- 2
sign.level <- 0.05
df.name <- 'Electricity'
df <- get(df.name)
spec <- formula('log(cost)~log(q)+I(log(q)^2)+log(pf)+log(pl)+log(pk)')

mod <- lm(formula=spec, data=df)
```

For the dataset ` `r df.name` ` consider a regression
**`r format(spec)`**. 

BP-test is based on the following heteroskedasicity model \(Var(u|X)=f(x'\gamma)\)

We test \(H_0:\gamma_1=\cdots=\gamma_k=0\) (homoskedastic error term)

Fit the regression and perform BP-test. Result is

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.bp.test <- bptest(mod)
my.bp.test
```

<span style="color: blue">significant level `r 100*sign.level`%</span>.

Evaluate necessary critical value  **Round the answer to `r my.digits` decimal places.**

```{r, echo=FALSE, comment=''}
chi2.cr <- qchisq(p=1-sign.level, df=my.bp.test$parameter[[1]] )
round(chi2.cr, digits = my.digits)
```

Conclusion

```{r echo=FALSE, comment=''}
if(my.bp.test$p.value[[1]]<sign.level) print('Evidence for heteroskedasticity') else print('Evidence for homoskedasticity')
```

## cost equation #2

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.digits.output <- 3
my.digits <- 2
sign.level <- 0.05
df.name <- 'Electricity'
df <- get(df.name)
spec <- formula('log(cost)~log(q)+I(log(q)^2)+log(pf)+log(pl)+log(pk)+log(pk)+I(log(pf)^2)+I(log(pl)^2)+I(log(pk)^2)')

mod <- lm(formula=spec, data=df)
```

For the dataset ` `r df.name` ` consider a regression
**`r format(spec)`**. 

BP-test is based on the following heteroskedasicity model \(Var(u|X)=f(x'\gamma)\)

We test \(H_0:\gamma_1=\cdots=\gamma_k=0\) (homoskedastic error term)

Fit the regression and perform BP-test. Result is

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.bp.test <- bptest(mod)
my.bp.test
```

<span style="color: blue">significant level `r 100*sign.level`%</span>.

Evaluate necessary critical value  **Round the answer to `r my.digits` decimal places.**

```{r, echo=FALSE, comment=''}
chi2.cr <- qchisq(p=1-sign.level, df=my.bp.test$parameter[[1]] )
round(chi2.cr, digits = my.digits)
```

Conclusion

```{r echo=FALSE, comment=''}
if(my.bp.test$p.value[[1]]<sign.level) print('Evidence for heteroskedasticity') else print('Evidence for homoskedasticity')
```

#  Robust inferences: t-test (HC s.e.)

By default we use HC3 s.e.

## output equation

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.digits.output <- 3
my.digits <- 2
sign.level <- 0.05
df.name <- 'Labour'
df <- get(df.name)
spec <- formula('log(output)~log(capital)+log(labour)+I(log(capital)^2)+I(log(labour)^2)')

mod <- lm(formula=spec, data=df)
```

For the dataset ` `r df.name` ` consider a regression
**`r format(spec)`**. 

Perform non-robust & robust t-test and compare results

Non-robust t-test (OLS-s.e.)

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
t.test <- coeftest(mod)
t.test
```

Robust t-test (HC3 s.e.)

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
robust.t.test <- coeftest(mod, vcov. = vcovHC)
robust.t.test
```

<span style="color: blue">significant level `r 100*sign.level`%</span>.

Which coefficients are significant?

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
regressors <- row.names(robust.t.test)
sign.regressors <- NULL
for(i in 1:nrow(robust.t.test) ) {
  if(robust.t.test[i,4]<sign.level) sign.regressors <- c(sign.regressors, 'Significant') else sign.regressors <- c(sign.regressors, 'Insignificant')
}
# names(sign.regressors) <- rownames(t.test)
print(data.frame(regressors, sign.regressors))
```

## cost equation

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.digits.output <- 3
my.digits <- 2
sign.level <- 0.01
df.name <- 'Electricity'
df <- get(df.name)
spec <- formula('log(cost)~log(q)+I(log(q)^2)+log(pf)+log(pl)+log(pk)')

mod <- lm(formula=spec, data=df)
```

For the dataset ` `r df.name` ` consider a regression
**`r format(spec)`**. 

Perform non-robust & robust t-test and compare results

Non-robust t-test (OLS-s.e.)

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
t.test <- coeftest(mod)
t.test
```

Robust t-test (HC3 s.e.)

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
robust.t.test <- coeftest(mod, vcov. = vcovHC)
robust.t.test
```

<span style="color: blue">significant level `r 100*sign.level`%</span>.

Which coefficients are significant?

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
regressors <- row.names(robust.t.test)
sign.regressors <- NULL
for(i in 1:nrow(robust.t.test) ) {
  if(robust.t.test[i,4]<sign.level) sign.regressors <- c(sign.regressors, 'Significant') else sign.regressors <- c(sign.regressors, 'Insignificant')
}
# names(sign.regressors) <- rownames(t.test)
print(data.frame(regressors, sign.regressors))
```

#   Robust inferences: F-test (HC estimator for the cov matrix)

By default we use HC3 estimator for covariance matrix

## output equation

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.digits.output <- 3
my.digits <- 2
sign.level <- 0.01
df.name <- 'Labour'
df <- get(df.name)
spec <- formula('log(output)~log(capital)+log(labour)+I(log(capital)^2)+I(log(labour)^2)')

mod <- lm(formula=spec, data=df)
```

For the dataset ` `r df.name` ` consider a regression **`r format(spec)`**. 

Robust t-test

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
robust.t.test <- coeftest(mod, vcov. = vcovHC)
robust.t.test
```
<span style="color: blue">significant level `r 100*sign.level`%</span>.

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
regressor <- 'labour'
```

Let's test the significance of **`r regressor`**, i.e. \(H_0:\beta_{\log(labour)}=\beta_{\log^2(labour)}=0\)

Result of non-robust F-test:
```{r,comment='', echo=FALSE}
testing.hypothesis <- c('log(labour)=0', 'I(log(capital)^2)=0')
F.test <- linearHypothesis(mod, testing.hypothesis )
# testing.result
stargazer(F.test[2,5:6], summary = FALSE, type = 'text', rownames=FALSE)
```

Result of robust F-test:
```{r,comment='', echo=FALSE}
robust.F.test <- linearHypothesis(mod, testing.hypothesis, white.adjust='hc3')
# testing.result
stargazer(robust.F.test[2,3:4], summary = FALSE, type = 'text', rownames=FALSE)
```

Evaluate necessary critical value. **Round the answer to `r my.digits` decimal places.**

```{r, echo=FALSE, comment=''}
F.cr <- qf(p=1-sign.level, df1 = robust.F.test[2,2], df2 = df.residual(mod))
round(F.cr, digits = my.digits)
```

Is **`r regressor`** significant?

```{r,comment='', echo=FALSE}
if(robust.F.test[2,4]<sign.level) print('Significant') else print('Insignificant')
```

## cost equation

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
my.digits.output <- 3
my.digits <- 2
sign.level <- 0.05
df.name <- 'Electricity'
df <- get(df.name)
spec <- formula('log(cost)~log(q)+I(log(q)^2)+log(pf)+log(pl)+log(pk)+log(pk)+I(log(pf)^2)+I(log(pl)^2)+I(log(pk)^2)')

mod <- lm(formula=spec, data=df)
```

For the dataset ` `r df.name` ` consider a regression
**`r format(spec)`**. 

Robust t-test

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
robust.t.test <- coeftest(mod, vcov. = vcovHC)
robust.t.test
```
<span style="color: blue">significant level `r 100*sign.level`%</span>.

### Hypothesis 1

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
regressor <- 'pl'
```

Let's test the significance of **`r regressor`**, i.e. \(H_0:\beta_{\log(pl)}=\beta_{\log^2(pl)}=0\)

Result of non-robust F-test:
```{r,comment='', echo=FALSE}
testing.hypothesis <- c('log(pl)=0', 'I(log(pl)^2)=0')
F.test <- linearHypothesis(mod, testing.hypothesis )
# testing.result
stargazer(F.test[2,5:6], summary = FALSE, type = 'text', rownames=FALSE)
```

Result of robust F-test:
```{r,comment='', echo=FALSE}
robust.F.test <- linearHypothesis(mod, testing.hypothesis, vcov. = vcovHC) #white.adjust='hc3')
# testing.result
stargazer(robust.F.test[2,3:4], summary = FALSE, type = 'text', rownames=FALSE)
```

Evaluate necessary critical value. **Round the answer to `r my.digits` decimal places.**

```{r, echo=FALSE, comment=''}
F.cr <- qf(p=1-sign.level, df1 = robust.F.test[2,2], df2 = df.residual(mod))
round(F.cr, digits = my.digits)
```

Is **`r regressor`** significant?

```{r,comment='', echo=FALSE}
if(robust.F.test[2,4]<sign.level) print('Significant') else print('Insignificant')
```

### Hypothesis 2

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
regressor <- 'pl'
```

Let's test the significance of **`r regressor`**, i.e. \(H_0:\beta_{\log(pk)}=\beta_{\log^2(pk)}=0\)

Result of non-robust F-test:
```{r,comment='', echo=FALSE}
testing.hypothesis <- c('log(pk)=0', 'I(log(pk)^2)=0')
# testing.hypothesis <- c('log(pl)=0', 'log(pk)=0')
F.test <- linearHypothesis(mod,  testing.hypothesis )
# testing.result
stargazer(F.test[2,5:6], summary = FALSE, type = 'text', rownames=FALSE)
```

Result of robust F-test:
```{r,comment='', echo=FALSE}
robust.F.test <- linearHypothesis(mod, testing.hypothesis, vcov. = vcovHC) #white.adjust='hc3')
# testing.result
stargazer(robust.F.test[2,3:4], summary = FALSE, type = 'text', rownames=FALSE)
```

Evaluate necessary critical value. **Round the answer to `r my.digits` decimal places.**

```{r, echo=FALSE, comment=''}
F.cr <- qf(p=1-sign.level, df1 = robust.F.test[2,2], df2 = df.residual(mod))
round(F.cr, digits = my.digits)
```

Is **`r regressor`** significant?

```{r,comment='', echo=FALSE}
if(robust.F.test[2,4]<sign.level) print('Significant') else print('Insignificant')
```

### Hypothesis 3

```{r, echo=FALSE, message=FALSE, warning=FALSE, comment=''}
regressor <- 'pl'
```

Let's test the significance of **`r regressor`**, i.e. \(H_0:\beta_{\log(pf)}=\beta_{\log^2(pf)}=0\)

Result of non-robust F-test:
```{r,comment='', echo=FALSE}
testing.hypothesis <- c('log(pf)=0', 'I(log(pf)^2)=0')
# testing.hypothesis <- c('log(pl)=0', 'log(pk)=0')
F.test <- linearHypothesis(mod,  testing.hypothesis )
# testing.result
stargazer(F.test[2,5:6], summary = FALSE, type = 'text', rownames=FALSE)
```

Result of robust F-test:
```{r,comment='', echo=FALSE}
robust.F.test <- linearHypothesis(mod, testing.hypothesis, vcov. = vcovHC) #white.adjust='hc3')
# testing.result
stargazer(robust.F.test[2,3:4], summary = FALSE, type = 'text', rownames=FALSE)
```

Evaluate necessary critical value. **Round the answer to `r my.digits` decimal places.**

```{r, echo=FALSE, comment=''}
F.cr <- qf(p=1-sign.level, df1 = robust.F.test[2,2], df2 = df.residual(mod))
round(F.cr, digits = my.digits)
```

Is **`r regressor`** significant?

```{r,comment='', echo=FALSE}
if(robust.F.test[2,4]<sign.level) print('Significant') else print('Insignificant')
```